All the concepts behind TrueSkill Through Time: the \proglang{Julia}, \proglang{Python} and \proglang{R} packages (Draft)

### 2020-11-06 article.tex
Articulo gaming: un poquito de modificaci'on en la intro

\en{Humans acquire and develop complex skills over the course of a lifetime.}
\es{Los humanos desarrollan habilidades complejas a lo largo de su vida.}
%
\en{This is possible by a combination of different biological, cognitive and social processes~\citep{Koster2020}.}
\es{Esto es posible gracias a una integraci\'on especial de los procesos biol\'ogicos, cognitivos y sociales~\citep{Koster2020}.}
%
\en{Our specie is characterized by first developing the brain and then the body to allow a longer period of learning~\citep{GonzalezForero2018}.}
\es{Nuestros cerebros se desarrollan primero, seguidos por el cuerpo porque esta secuencia permite un per\'iodo m\'as largo de aprendizaje~\citep{GonzalezForero2018}.}
%
% \en{Human life histories are distinguished by long periods of juvenile dependency, short intervals between births, and long postreproductive life span~\citep{Jones2011}.}
% \es{Las historias de vida humanas se distinguen por largos per\'iodos de dependencia juvenil, cortos intervalos entre nacimientos y una larga vida posreproductiva~\citep{Jones2011}.}
%
\en{A special cognitive ability allows humans to learn things from others, and transmit innovations through generations~\citep{Herrmann2007}.}
\es{Una capacidad cognitiva especial le permite a los humanos aprender de los dem\'as y transmitir las innovaci\'on a través de la generaciones~\citep{Herrmann2007}.}
%
\en{As a population-based process, human adaptation is also affected by demographic characteristics, such as the size and structure of populations~\citep{derex2020-populationStructure}.}
\es{Al ser un proceso poblacional, la adaptaci\'on humana también se ve afectada por caracter\'isticas demogr\'aficas, como el tama\~no y estructura de las poblaciones~\citep{derex2020-populationStructure}.}

@@

\en{Understanding the factors impacting in the positive or negative changes of skill could be used to develop new training strategies or adapting the environment to avoid unnecessary tasks.}
\es{Conocer c\'omo cambian la habilidades de la personas en el tiempo es crucial en muchas \'areas.}
%
\en{For example, in sports development and education, knowing the skill and how it changes can serve to avoid exposing agents to imbalanced challenges, maintain the interest and engagement.}
\es{En contextos educativos, calificaciones precisas pueden servir para evitar exponer a los agentes a desaf\'ios desproporcionados, estimulando as\'i el inter\'es y la participaci\'on.}
%
\en{For research purposes, they can help to refine and test models of human learning.}
\es{Con fines de investigaci\'on, ellas pueden ayudar a refinar y testear modelos de aprendizaje humano.}
%
\en{In practice, since skill is a hidden variable, the best we can do is estimate it from its direct observable consequences.}
\es{En la pr\'actica, dado que la habilidad es una variables oculta, lo mejor que podemos hacer es estimarla a partir de sus consecuencias observables m\'as directas.}
%
\en{The most widely used sources of information in education and sports are the outcomes of problem solving and competition between opponents.}
\es{Las fuentes de informaci\'on m\'as utilizadas en educaci\'on y deportes provienen del producto de resoluci\'on de problemas y de competencia entre oponentes.}
%
\en{The naive solution, frequency of positive results, is not a good indicator essentially because it also depends on the difficulty of the challenges.}
\es{La soluci\'on ingenua, frecuencia de resultados positivos, no es un buen indicador fundamentalmente porque depende tambi\'en de la dificultad de los desaf\'ios.}
%
\en{For that reason, all solutions rely on pairwise comparisons.}
\es{Por esa raz\'on, todos las soluciones se basan en comparaciones entre pares.}
%
\en{Since the first generative models, proposed almost a century ago by~\cite{thurstone1927-comparativeJugement} and~\cite{zermelo1929-translation}, it is assumed that the probability of an observed result, $r_{ij}$, depends on the difference in performance, $d_{ij}$, between the agent $i$ and target $j$, $P(\, r_{ij} \,|\, d_{ij} \,)$.}
\es{Ya los primeros modelos generativos, propuestos hace casi un siglo por~\cite{thurstone1927-comparativeJugement} y~\cite{zermelo1929-translation}, suponen que la probabilidad de un resultado observado, $r$, depende de la diferencia de rendimiento, $d$, entre el agente y su objetivo, $P(\, r_{ij} \,|\, d_{ij} \,)$.}
%
\en{The field revitalized 20 years later by the publications of~\cite{bradley1952-pairedComparisions} and~\cite{mosteller1951-remarks}.}
\es{El campo revitaliz\'o 20 a\~nos despu\'es gracias a las publicaciones de \cite{bradley1952-pairedComparisions} y~\cite{mosteller1951-remarks}.}

>>>>

\en{Humans acquire and develop complex skills over the course of a lifetime.}
\es{Los humanos desarrollan habilidades complejas a lo largo de su vida.}
%
\en{This is possible by a combination of different biological, cognitive and social processes~\citep{Koster2020}.}
\es{Esto es posible gracias a una integraci\'on especial de los procesos biol\'ogicos, cognitivos y sociales~\citep{Koster2020}.}
%
\en{Our specie is characterized by first developing the brain and then the body to allow a longer period of learning~\citep{GonzalezForero2018}.}
\es{Nuestros cerebros se desarrollan primero, seguidos por el cuerpo porque esta secuencia permite un per\'iodo m\'as largo de aprendizaje~\citep{GonzalezForero2018}.}
%
% \en{Human life histories are distinguished by long periods of juvenile dependency, short intervals between births, and long postreproductive life span~\citep{Jones2011}.}
% \es{Las historias de vida humanas se distinguen por largos per\'iodos de dependencia juvenil, cortos intervalos entre nacimientos y una larga vida posreproductiva~\citep{Jones2011}.}
%
\en{A special cognitive ability allows humans to learn things from others, and transmit innovations through generations~\citep{Herrmann2007}.}
\es{Una capacidad cognitiva especial le permite a los humanos aprender de los dem\'as y transmitir las innovaci\'on a través de la generaciones~\citep{Herrmann2007}.}
%
\en{As a population-based process, human adaptation is also affected by demographic characteristics, such as the size and structure of populations~\citep{Derex2020}.}
\es{Al ser un proceso poblacional, la adaptaci\'on humana también se ve afectada por caracter\'isticas demogr\'aficas, como el tama\~no y estructura de las poblaciones~\citep{Derex2020}.}

@@

\en{Understanding the factors impacting in the positive or negative changes of skill could be used to develop new training strategies or adapting the environment to avoid unnecessary tasks.}
\es{Conocer c\'omo cambian la habilidades de la personas en el tiempo es crucial en muchas \'areas.}
%
\en{For example, in sports development and education, knowing the skill and how it changes can support avoiding exposing agents to imbalanced challenges, maintaining the interest and engagement, and monitoring the learning and improvement process.}
\es{En contextos educativos, calificaciones precisas pueden servir para evitar exponer a los agentes a desaf\'ios desproporcionados, estimulando as\'i el inter\'es y la participaci\'on.}
%
\en{In practice, since skill is a hidden variable, the best we can do is estimating it from its direct observable consequences.}
\es{En la pr\'actica, dado que la habilidad es una variables oculta, lo mejor que podemos hacer es estimarla a partir de sus consecuencias observables m\'as directas.}
%
\en{The most widely used sources of information in education and sports are the outcomes of problem solving and competition between opponents.}
\es{Las fuentes de informaci\'on m\'as utilizadas en educaci\'on y deportes provienen del producto de resoluci\'on de problemas y de competencia entre oponentes.}
%
\en{The naive solution (i.e. frequency of positive results) is not a good estimator because it depends on the difficulty of the challenges.}
\es{La soluci\'on ingenua, frecuencia de resultados positivos, no es un buen indicador fundamentalmente porque depende tambi\'en de la dificultad de los desaf\'ios.}
%
\en{The estimation should rely on pairwise comparisons.}
\es{Por esa raz\'on, todos las soluciones se basan en comparaciones entre pares.}
%
\en{Since the first generative models, proposed almost a century ago by~\cite{thurstone1927-comparativeJugement} and~\cite{zermelo1929-translation}, it is assumed that the probability of an observed result, $r_{ij}$, depends on the difference in performance, $d_{ij}$, between the agent $i$ and target $j$, $P(\, r_{ij} \,|\, d_{ij} \,)$.}
\es{Ya los primeros modelos generativos, propuestos hace casi un siglo por~\cite{thurstone1927-comparativeJugement} y~\cite{zermelo1929-translation}, suponen que la probabilidad de un resultado observado, $r$, depende de la diferencia de rendimiento, $d$, entre el agente y su objetivo, $P(\, r_{ij} \,|\, d_{ij} \,)$.}
%
\en{The field revitalized 20 years later by the publications of~\cite{bradley1952-pairedComparisions} and~\cite{mosteller1951-remarks}.}
\es{El campo revitaliz\'o 20 a\~nos despu\'es gracias a las publicaciones de \cite{bradley1952-pairedComparisions} y~\cite{mosteller1951-remarks}.}
